[![Build Status](https://travis-ci.org/mpadge/hotspotr.svg?branch=master)](https://travis-ci.org/mpadge/hotspotr)
[![codecov](https://codecov.io/gh/mpadge/hotspotr/branch/master/graph/badge.svg)](https://codecov.io/gh/mpadge/hotspotr)

R package for estimating whether the statistical properties of a spatial pattern
of hotspots may be reproduced with a simple neutral model. Currently only works
for gridded data.

## Install

```{r, eval=FALSE}
devtools::install_github ('mpadge/hotspotr')
```
```{r load, echo=FALSE, message=FALSE}
while (length (grep ('hotspotr', getwd ())) > 0) setwd ("..")
#devtools::document ("hotspotr")
#devtools::check ("hotspotr")
devtools::load_all ("hotspotr")
setwd ("./hotspotr")
```

## Test

First a demonstration with a seed that produces a 2D field very similar to
neutral fields.

```{r demo, fig.width=12, fig.height=6}
seed <- 18
ydat <- ives2D (size=10, seed=seed)
x11 ()
par (mfrow=c(1,2))
test <- test1d (ydat, plot=TRUE)
test <- test2d (ydat, plot=TRUE)
```

Then the text output of `run_tests` for a random seed giving more typically low
p-values
```{r run_tests}
run_tests (ntests=1000)
```

